---
title: "Naive Bayes"
output: html_document
---

Loading the Data and installing required packages & Libraries

```{r}

Mydata <- read.csv("UniversalBank.csv")
#summary(Mydata)

#install.packages("dplyr")
library(dplyr)
#install.packages("caret")
library(caret)
#install.packages("caTools")
library(caTools)
# install.packages("e1071")
library(e1071)

```

***

Factoring the required variables (because ftable function takes factored variables) and Dividing the dataset into 60% training and 40% validation sets

However, even without factoring the algorithim would produce the same results, perhaps, as all the values are already kind of factored with 1s and 0s.

```{r}

Mydata$Personal.Loan = factor(Mydata$Personal.Loan)
Mydata$Online = factor(Mydata$Online)
Mydata$CreditCard = factor(Mydata$CreditCard)

set.seed(1)

split = sample.split(Mydata$Personal.Loan,SplitRatio = 0.6)
training_set = subset(Mydata, split == TRUE)
val_set = subset(Mydata, split == FALSE)

```

***

## Questions:

A. Create a pivot table for the training data with Online as a column variable, CC as a row
variable, and Loan as a secondary row variable. 

```{r}


ftable(training_set$CreditCard,training_set$Personal.Loan,
      training_set$Online, dnn = c("CreditCard", "Personal.Loan", "Online"))


```

***

B. Consider the task of classifying a customer who owns a bank credit card and is actively using
online banking services. Looking at the pivot table, what is the probability that this customer
will accept the loan offer? [This is the probability of loan acceptance (Loan = 1) conditional on
having a bank credit card (CC = 1) and being an active user of online banking services (Online
= 1)]



Lets say, P(Loan1) = P(L1), P(Loan0)= p(L0) , P(CC1)= P(CC1) and P(Online1) = P(O1)


The solution can be written as P(L1)|P(CC1) , P(O1)

Further, it can be simplified as , (P(L1) and P(CC1) and P(O1))/P(CC1) and P(O1)

From A , the numerator will be 49 and the denominator will be 531 (482+49)

Hence, solution will be = 49/531 or 9.22%

***



C. Create two separate pivot tables for the training data. One will have Loan (rows) as a
function of Online (columns) and the other will have Loan (rows) as a function of CC.

```{r}

ftable(training_set$Personal.Loan,training_set$Online, dnn = c("Personal.Loan", "Online"))
ftable(training_set$Personal.Loan,training_set$CreditCard, dnn = c("Personal.Loan", "CreditCard"))


```
***

D. Compute the following quantities [P(A | B) means “the probability ofA given B”]:
i. P(CC = 1 | Loan = 1) (the proportion of credit card holders among the loan acceptors)
ii. P(Online = 1 | Loan = 1)
iii. P(Loan = 1) (the proportion of loan acceptors)
iv. P(CC = 1 | Loan = 0)
v. P(Online = 1 | Loan = 0)
vi. P(Loan = 0)

## Answers:

Lets say, P(Loan1) = P(L1), P(Loan0)= p(L0) , P(CC1)= P(CC1) and P(Online1) = P(O1)

i.  Can be written as P(L1 and CC1)/P(L1) , so it would be , 92/(196+92) = 92/288 => 31.94%
ii. Can be written as P(L1 and O1)/P(L1) , so it would be , 168/(120+168) = 168/288 => 58.33%
iii.288/Total Number ===> 288/3000  ==> 9.6%
iv. Can be written as P(L0 and CC1)/P(L0) , so it would be , 794/(1918+794) = 794/2712  => 29.27%
v.  Can be written as P(L0 and O1)/P(L0) , so it would be , 1627/(1085+1627) = 1627/2712 => 59.99%
vi. 2712/Total Number ===> 2712/3000 ==> 90.4%

***

E. Use the quantities computed above to compute the naive Bayes probability P(Loan = 1 | CC
= 1, Online = 1).


Lets say, P(Loan1) = P(L1), P(Loan0)= p(L0) , P(CC1)= P(CC1) and P(Online1) = P(O1)


The solution can be written as P(L1)|P(CC1) , P(O1)


Formula for the same using Naive Bayes theorm is P(CC1)|P(L1) * P(O1)|P(L1) * P(L1) / (P(CC1 and O1))

Assumption: Naive Bayes theorm assumes that CC1 and O1 are independent given L1

Can be further written as P(CC1)|P(L1) * P(O1)|P(L1) * P(L1) / ( (P(CC1)|P(L1) * P(O1)|P(L1) * P(L1) ) + (P(CC1)|P(L0) * P(O1)|P(L0) * P(L0) ) )

Thus (we can get the values from " D" ) , = (92/288) * (168/288) * (288/3000) / (((92/288) * (168/288) * (288/3000)) + ((794/2712) * (1627/2712) * (2712/3000)))

That will be ==> 10.12%


***

F.Compare this value with the one obtained from the pivot table in (B). Which is a more
accurate estimate?


Value from B is 9.22 and E is 10.12.

Naive Bayes theorm assumes that CC1 and O1 are independent of each other given L1 . But in reality there may be several factors making an impact on each other and can be dependent ( we dont know). Thus naive Bayes method may not be as accurate as the direct method. 

Hence value , which is 9.22,  from B may be more accurate.

***


G.Which of the entries in this table are needed for computing P(Loan = 1 | CC = 1, Online = 1)?
Run naive Bayes on the data. Examine the model output on training data, and find the entry
that corresponds to P(Loan = 1 | CC = 1, Online = 1). Compare this to the number you
obtained in (E).


```{r}
model1 <- naiveBayes(Personal.Loan ~ ., data = training_set[c(10,13,14)])
model1
```

***

1. From the table we need only Index 13 and 14 (Online and Credit card). Hence using those two indexes for the model building.
2. Values from the two predictors exactly match with what we have got in E and/or D (.3194, .5833 , .2927 and .5999 ) and P(L1) which is 0.096 and P(L0) which is 0.904 exactly match with A-priori probabilities.




    
